Blum axioms and nondeterministic computation of functions

### 2016-05-18 paper.tex
paper.tex: Merge sections "Our approach" and "Quality standards".

This paper presents several attempts to establish
how a nondeterministic Turing machine could compute a function,
and to extend the concept of ``polynomial-time computable'' under each definition.
Section~\ref{sec:quality-standards} sets reasonableness criteria
to both the definition of function computation and complexity of the computation.
Section~\ref{sec:nondeterministic_computation_of_functions},
which contains the several definitions,
presents and criticizes Hopcroft and Ullman's and Goldreich's definitions
(sections \ref{sec:hopcroft-ullman} and~\ref{sec:goldreich}),
proposes one possible definition that meet the quality standards
(section~\ref{sec:proposed}),
and shows other two definitions,
by Krentel (sec~\ref{sec:krentel}) and Valiant (sec~\ref{sec:valiant}),
that, altough were not proposed in the context
of general nondeterministic computation of functions,
also meet the quality standards and sidesteps the problems with the proposed definition
(presented in section~\ref{sec:np-completeness-extension}).
The development of this project is mentioned in section~\ref{sec:development},
after the concluding remarks (section~\ref{sec:conclusion}).

\section{Our approach}
\label{sec:our-approach}

The mathematical definition of function
requires it to return a value for all elements of its domain,
and to return a single value in all cases.
In this paper,
we will try to associate nondeterministic computation
with \emph{partial} recursive functions;
that is,
functions that are only defined for some elements in its domain.
(Strictly speaking,
these objects are not ``functions'';
but we will allow for this abuse of nomenclature.
If the partial function is defined everywhere,
we will call it a \emph{total} function, for contrast.)
In general,
we cannot demand a function resulting from a computation to be total
because the computation may not halt.

@@

\section{Quality Standards}
\label{sec:quality-standards}

We are trying to make a definition of non-deterministic computation of functions.
We want our definition to be reasonable,
in the sense of not falling in the realm of undecidable problems,
nor ``forgetting'' some computable function;
and we want our definition to be efficient,
in the sense to preserve the apparent\footnote{
    It is apparent in the sense that,
    if proven would show $\P \neq \NP$,
    and if disproved (showing a polynomial slowdown is the best we can do)
    would show $\P = \NP$.
    Although most researchers expect the former to be the case~\cite[p.~54]{Gasarch2012},
    as Papadimitriou noted~\cite[p.~412]{Papadimitriou1994},
    in the absence of a proof that $\P \neq \NP$
    we should not be too emphatic in stating the simulation \emph{require}
    (as opposed to \emph{seemingly require}) exponential slowdown.
}
exponential speed-up when solving,
for instance, the Boolean satisfiability problem
--- that is, there should be an exponential speed-up
when computing the characteristic function\footnotemark{} of $\SAT$.
\footnotetext{
    The characteristic function of a set $A$
    is the function $1_A$ defined to be $1$ for $x \in A$
    and $0$ for $x \notin A$.
}

Sections \ref{sec:godel-numberings} and~\ref{sec:blum-axioms}
formalize these reasonableness and efficiency requirements,
respectively.

@@

Sections \ref{sec:godel-numberings} and~\ref{sec:blum-axioms}
introduced the concept of acceptable Gödel numbering,
the Blum axioms,
and its complexity classes.
As we want to regard these concepts as ``quality requirements''
for the definitions,
we will analyze each of them to see whether they define acceptable Gödel numberings,
the analogous time complexity satisfies Blum axioms,
and that the characteristic function of the Boolean satisfiability problem
can be solved in ``polynomial time'' according to that complexity measure.

>>>>

This paper presents several attempts to establish
how a nondeterministic Turing machine could compute a function,
and to extend the concept of ``polynomial-time computable'' under each definition.
Section~\ref{sec:our-approach} sets reasonableness criteria
to both the definition of function computation and complexity of the computation.
Section~\ref{sec:nondeterministic_computation_of_functions},
which contains the several definitions,
presents and criticizes Hopcroft and Ullman's and Goldreich's definitions
(sections \ref{sec:hopcroft-ullman} and~\ref{sec:goldreich}),
proposes one possible definition that meet the quality standards
(section~\ref{sec:proposed}),
and shows other two definitions,
by Krentel (sec~\ref{sec:krentel}) and Valiant (sec~\ref{sec:valiant}),
that, altough were not proposed in the context
of general nondeterministic computation of functions,
also meet the quality standards and sidesteps the problems with the proposed definition
(presented in section~\ref{sec:np-completeness-extension}).
The development of this project is mentioned in section~\ref{sec:development},
after the concluding remarks (section~\ref{sec:conclusion}).

\section{Our approach: Blum axioms and Gödel numberings}
\label{sec:our-approach}

In this paper,
we will try to associate nondeterministic computation
with partial recursive functions, in some well-behaved manner,
and preserving the apparent\footnotemark{} exponential speed-up
present in nondeterministic deciders.
\footnotetext{
    It is apparent in the sense that,
    if proven would show $\P \neq \NP$,
    and if disproved (showing a polynomial slowdown is the best we can do)
    would show $\P = \NP$.
    Although most researchers expect the former to be the case~\cite[p.~54]{Gasarch2012},
    as Papadimitriou noted~\cite[p.~412]{Papadimitriou1994},
    in the absence of a proof that $\P \neq \NP$
    we should not be too emphatic in stating the simulation \emph{require}
    (as opposed to \emph{seemingly require}) exponential slowdown.
}
The concept of Gödel numberings
(section~\ref{sec:godel-numberings})
captures the notion of ``well-behaved''.
The Blum axioms
(section~\ref{sec:blum-axioms})
capture the notions of computational complexity.
We thus will demand the definitions to satisfy
the requirements of Gödel numberings and Blum axioms.

@@

@@

Sections \ref{sec:godel-numberings} and~\ref{sec:blum-axioms}
introduced the concept of acceptable Gödel numbering,
the Blum axioms,
and its complexity classes.
As we want to regard these concepts as ``quality requirements''
for the definitions,
we will analyze each of them to see whether they define acceptable Gödel numberings,
the analogous time complexity satisfies Blum axioms,
and that the characteristic function\footnotemark{} of the Boolean satisfiability problem
can be solved in ``polynomial time'' according to that complexity measure.
\footnotetext{
    The characteristic function of a set $A$
    is the function $1_A$ defined to be $1$ for $x \in A$
    and $0$ for $x \notin A$.
}
(This last one is an indicative that
the definition preserves the exponential speed-up that nondeterminism gives to deciders.)
